Coalescent inference of mutation spectrum histories from sample frequency spectra

### 2019-09-03 manuscript/main.tex
changed writing that explains the cost function

\usepackage{fullpage}
\usepackage[round,semicolon]{natbib}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{hyperref}

@@

\author{
W. DeWitt$^{1}$, K. Harris$^{2}$, and K. Harris$^{1}$\\
\small{Departments of $^1$Genome Sciences and $^2$Biology, University of Washington, Seattle, USA}
% \small{$^\ast$ Equal contribution}
}

@@

%WD more inverse problem intro
The inverse problem \eqref{eqn:transform} is ill-posed in general, so many very different histories can be equally consistent with the data \citep{oscillation paper? Yun's other papers?}.
To penalize oscillatory histories we regularize on the derivative of $\mu$.
We define $L_p^p$ derivative regularizers for history $\mu(t)$:
\[
\int_0^\infty\left|\frac{d\mu(t)}{dt}\right|^p dt.
\]
In the finite-dimensional approximation \eqref{eqn:transform_discrete} the derivatives become Dirac delta functions at the change points, and this become
\[
\left|\Delta\boldsymbol z\right|_p^p.
\]
where $\Delta$ denotes the first difference operator.
Our penalized log-likelihood is then
\begin{equation}
\label{eqn:penalized}
\tilde\ell(\boldsymbol z) = \ell(\boldsymbol z) - \lambda \left(\alpha\left|\Delta\boldsymbol z\right|_1 + \frac{(1-\alpha)}{2}\left|\Delta\boldsymbol z\right|_2^2\right),
\end{equation}
where $\lambda$ tunes the regularization strength, and $\alpha$ tunes the relative $p=1$ vs $p=2$ regularization.

>>>>

\usepackage{fullpage}
\usepackage[round,semicolon]{natbib}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{color}
\usepackage{hyperref}

@@

\author{
W. DeWitt$^{1}$, K. Decker Harris$^{2,3}$, and K. Harris$^{1}$\\
{\small
$^1$Genome Sciences,
$^2$Paul G.\ Allen School of Computer Science \& Engineering, 
and $^3$Biology}\\
{\small University of Washington, Seattle, USA}
% \small{$^\ast$ Equal contribution}
}

@@

%WD more inverse problem intro
The inverse problem \eqref{eqn:transform} is ill-posed in general, so many very different histories can be equally consistent with the data \citep{oscillation paper? Yun's other papers?}.
In order to recover smooth histories, 
we penalize the $L^p$ norms
of the derivative of $\mu$ 
\[
\left\| \frac{d \mu}{d t} \right\|_p^p 
= \int_0^\infty\left|\frac{d\mu(t)}{dt}\right|^p dt,
\]
for $p=1$ and 2.
With time discretized \eqref{eqn:transform_discrete},
this becomes
\[
\left\|\Delta\boldsymbol z \right\|_p^p.
\]
where $\Delta$ denotes the first difference matrix.
Our cost function (penalized log-likelihood) is then
\begin{equation}
\label{eqn:penalized}
C(\bm z)
%\tilde\ell(\boldsymbol z) 
= -\ell(\boldsymbol z) + \lambda \left(\alpha\left\|\Delta\boldsymbol z\right\|_1 + \frac{(1-\alpha)}{2}\left\|\Delta\boldsymbol z\right\|_2^2\right),
\end{equation}
where $\lambda$ tunes the regularization strength, and $\alpha$ tunes the relative $p=1$ vs $p=2$ regularization.
\textcolor{blue}{Kam note: probably we need a ``Ridge'' (Tikhonov) 
term like the following to deal with small singular values:}
\[
\lambda' \| \bm z \|_2^2
\]
